---
title: "p8106_mtp"
author: "Abhishek Ajay and PengXuan Chen"
date: "April 2, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(glmnet)
library(caret)
library(corrplot)
library(splines)
library(ggplot2)
library(mgcv)
library(earth)
library(ggpubr)
library(knitr)
set.seed(2)
```


```{r}
day = read.csv("./data/day.csv") %>% 
  janitor::clean_names() %>% 
  select(-c(dteday, instant)) %>% 
  mutate(season = as.factor(season), 
         yr = as.factor(yr), 
         mnth = as.factor(mnth), 
         holiday = as.factor(holiday), 
         weekday = as.factor(weekday), 
         workingday = as.factor(workingday),
         weathersit = as.factor(weathersit)
         #dteday = as.vector(dteday)
         ) 
  
x = day  
x = model.matrix(casual + registered + cnt~., x)[,-1]
y = day$cnt
y_n = 0.01*(day$cnt)
corrplot(cor(x))
a = day %>% 
  select(-c(yr, mnth, weekday))
x_n = model.matrix(casual + registered + cnt~.,a)[,-1]

```


```{r}
#simple linear regression

model_lm = lm(cnt~ season +  holiday +  workingday + weathersit + temp + atemp + hum + windspeed, data = day)
summary(model_lm)
model_lm2 = lm(0.01*cnt~ season +  holiday +  workingday + weathersit + temp + atemp + hum + windspeed, data = day)
summary(model_lm)
ctrl1 <-trainControl(method = "cv", number = 10)
set.seed(1)
lmFit <-train(cnt~ season +   holiday  + workingday + weathersit + temp + atemp + hum + windspeed,data = day,method = "lm",trControl = ctrl1)
lm.mse = (lmFit$results[,2])^2
lmFit2 <-train(0.01*cnt~ season +   holiday  + workingday + weathersit + temp + atemp + hum + windspeed,data = day,method = "lm",trControl = ctrl1)
lm.mse2 = (lmFit2$results[,2])^2


```


mse  =1716100

```{r}
#Non Linear Methods from L5.pdf
theme1 <-trellis.par.get()
theme1$plot.symbol$col <-rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <-rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <-rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x_n, y, plot = "scatter", labels =c("","Y"),type =c("p"), layout =c(4, 3))
```

The above plots clearly show that there's non linear relationship between "cnt" and temp, atemp, hum and windspeed. What follows is us trying to fit univariate non-linear models and GAM to investigate the association between cnt and the four predictors.


smooth splines 
```{r}
#temp
fit.ss <-smooth.spline(day$temp, day$cnt)
fit.ss$df

templims <-range(day$temp)
temp.grid <-seq(from = templims[1],to = templims[2], by = 0.0001)
length(temp.grid)

pred.ss <-predict(fit.ss,
                  x = temp.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        temp = temp.grid)

p <-ggplot(data = day,aes(x = temp, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

ssp1 = p+geom_line(aes(x = temp, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()

#humidity
fit.ss <-smooth.spline(day$hum, day$cnt)
fit.ss$df

humlims <-range(day$hum)
hum.grid <-seq(from = humlims[1],to = humlims[2], by = 0.001)
length(hum.grid)

pred.ss <-predict(fit.ss,
                  x = hum.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        hum = hum.grid)

p <-ggplot(data = day,aes(x = hum, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

ssp3 = p+geom_line(aes(x = hum, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()

#pretty persuasive, since 0.5-0.75 is the best comfortable humidity

#windspeed
fit.ss <-smooth.spline(day$windspeed, day$cnt)
fit.ss$df

windspeedlims <-range(day$windspeed)
windspeed.grid <-seq(from = windspeedlims[1],to = windspeedlims[2], by = 0.001)
length(windspeed.grid)

pred.ss <-predict(fit.ss,
                  x = windspeed.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        windspeed = windspeed.grid)

p <-ggplot(data = day,aes(x = windspeed, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

ssp4 = p+geom_line(aes(x = windspeed, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()


# windspeed ^  --  cnt V

#atemp
fit.ss <-smooth.spline(day$atemp, day$cnt)
fit.ss$df

atemplims <-range(day$atemp)
atemp.grid <-seq(from = atemplims[1],to = atemplims[2], by = 0.01)
length(atemp.grid)

pred.ss <-predict(fit.ss,
                  x = atemp.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        atemp = atemp.grid)

p <-ggplot(data = day,aes(x = atemp, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

ssp2 = p+geom_line(aes(x = atemp, y = pred), data = pred.ss.df, color =rgb(.8, .1, .1, 1))+ theme_bw()

par(mfrow=c(2,2))
ggarrange(ssp1, ssp2,ssp3,ssp4,
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
```

##GAM
```{r}
gam.m1 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed, data = day)

gam.m2 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + s(temp) + atemp + hum + windspeed, data = day) # s:temp
gam.m3 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + s(atemp) + hum + windspeed, data = day) # s:atemp
gam.m4 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + s(hum) + windspeed, data = day) # s:hum
gam.m5 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + hum + s(windspeed), data = day) # s:windspeed

anova(gam.m1, gam.m2, test = "F")
anova(gam.m1, gam.m3, test = "F")  
anova(gam.m1, gam.m2, gam.m3, gam.m4, gam.m5, test = "F")

par(mfrow=c(2,2))
plot(gam.m2)
plot(gam.m3)
plot(gam.m4)
plot(gam.m5)

#deleting the yr and mnth and weekday variables

gam.m1a = gam(cnt~ season +  holiday  + workingday + weathersit + temp + atemp + hum + windspeed, data = day)


gam.m2a = gam(cnt~ season +  holiday  + workingday + weathersit + s(temp) + atemp + hum + windspeed, data = day) # s:temp
gam.m3a = gam(cnt~ season +  holiday  + workingday + weathersit + temp + s(atemp) + hum + windspeed, data = day) # s:atemp
gam.m4a = gam(cnt~ season +  holiday  + workingday + weathersit + temp + atemp + s(hum) + windspeed, data = day) # s:hum

gam.m5a = gam(cnt~ season +  holiday  + workingday + weathersit + temp + atemp + hum + s(windspeed), data = day) # s:windspeed

anova(gam.m1, gam.m2, test = "F")
anova(gam.m1, gam.m3, test = "F")

par(mfrow=c(2,2))
plot(gam.m2a)
plot(gam.m3a)
plot(gam.m4a)
plot(gam.m5a)


```

```{r}
#GCV to select df

ctrl1 <-trainControl(method = "cv", number = 10)
set.seed(2)
gam.fit <-train(x, y,method = "gam",tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),trControl = ctrl1)

gam.fit$bestTune
gam.fit$finalModel

x_2 = 
  day %>% 
  select(-c( casual, registered, cnt, yr, mnth, weekday))

x_2 = model.matrix(~., x_2)



set.seed(3)
gam.fit2 <-train(x_2, y,method = "gam",tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),trControl = ctrl1)

gam.fit2$bestTune
gam.mse = gam.fit2$finalModel$gcv.ubre

gam.fit3 <-train(x_2, y_n,method = "gam",tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),trControl = ctrl1)
gam.mse2 = gam.fit3$finalModel$gcv.ubre
```

?The deviance is pretty huge of those ANOVA?

GCV score is 1418371 for gam.fit2. 

```{r}
#poisson regression for categorical terms, exclude yr and mnth
day_p = day %>% 
  mutate(cnt = as.integer(0.01*cnt))
day.glm1 <- glm(cnt~season +  holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed, family=poisson, data=day)
summary(day.glm1)
pchisq(day.glm1$deviance, df=day.glm1$df.residual, lower.tail=FALSE)

day.glm2 <- glm(cnt~season +  holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed, family=poisson, data=day_p)
summary(day.glm2)
pchisq(day.glm2$deviance, df=day.glm2$df.residual, lower.tail=FALSE)

exp(day.glm1$coefficients)
```

poisson regression is a nonlinear regression for discrete outcomes(when response is a count).  However, it is only valid when large outcome is not frequent in the dataset. 
the variables obtaining collinearity with each other potentially will results in an "NA" in the model, thus we chose to eliminate yr and mnth. workingday also seems to have perfect collinearity with weekday&holiday.
The large scale of our outcome has leads to an unreasonable high deviance in our model. By learning how bad the goodness of fit is, we tried to shrink the outcome by mutiplying 0.01 to it. This gives us similar estimates however with a much better and smaller deviance, which results in a pearson chi-sq test with a p-value that is close to zero. 
**SKIP this**

```{r}
#mars
mars_grid <-expand.grid(degree = 1:2,nprune = 2:30)
set.seed(2)
mars.fit <-train(x_n, y,method = "earth",tuneGrid = mars_grid,trControl = ctrl1)

ggplot(mars.fit)
mars.fit$bestTune
mars.mse = mars.fit$finalModel$gcv
summary(mars.fit)
coef(mars.fit$finalModel)
bwplot(resamples(list(mars = mars.fit,gam = gam.fit2)), metric = "RMSE")

mars.fit2 <-train(x_n, y_n,method = "earth",tuneGrid = mars_grid,trControl = ctrl1)
mars.mse2 = mars.fit2$finalModel$gcv
```

GCV 1414719

#Linear Model (less flexibility but more interpretability)


```{r}


ridge.mod <-glmnet(x_n, y, alpha=0, lambda =exp(seq(-1, 10, length=100)))
ridge.mod$lambda

dim(coef(ridge.mod))
```

we dropped the dteday because it does is a factor variable with too many levels and it basically is the same as the index which does not tells us anything in this test. 

x_1: all independent variables
y: cnt---the response

```{r}
set.seed(2)
cv.ridge <-cv.glmnet(x_n, y,alpha = 0,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")
plot(cv.ridge)

plot_glmnet(ridge.mod, xvar = "rlambda")

best.lambda = cv.ridge$lambda.min
best.lambda
predict(cv.ridge, s = "lambda.min", type="coefficients")
ridge.mse = min(cv.ridge$cvm)


cv.ridge2 <-cv.glmnet(x_n, y_n,alpha = 0,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")



predict(cv.ridge2, s = "lambda.min", type="coefficients")
ridge.mse2 = min(cv.ridge2$cvm)
```

the min mse is `r min(cv.ridge$cvm)`. 

```{r}
cv.lasso <-cv.glmnet(x_n, y,alpha = 1,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")
cv.lasso$lambda.min
plot(cv.lasso)

plot_glmnet(cv.lasso$glmnet.fit)

predict(cv.lasso, s="lambda.min", type="coefficients")

lasso.mse = min(cv.lasso$cvm)

cv.lasso2 <-cv.glmnet(x_n, y_n,alpha = 1,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")


predict(cv.lasso2, s="lambda.min", type="coefficients")

lasso.mse2 = min(cv.lasso2$cvm)

```

? why is that lasso best lambda is 1 ?
the min mse is `r min(cv.lasso$cvm)`


```{r}
#table mse



table1 = matrix(cbind(mars.mse,gam.mse,ridge.mse,lm.mse,lasso.mse, mars.mse2,gam.mse2,ridge.mse2,lm.mse2,lasso.mse2 ), ncol = 2)
colnames(table1) = c("MSE","MSE for scaled outcome")
rownames(table1) = c("MARS", "GAM", "ridge","Linear","LASSO")
table1 = knitr::kable(table1)
table1


#table coef


as.numeric(coef(model_lm))
as.numeric(coef(model_lm2))

as.numeric(predict(cv.ridge, s = "lambda.min", type="coefficients"))
as.numeric(predict(cv.ridge2, s = "lambda.min", type="coefficients"))

as.numeric(predict(cv.lasso, s = "lambda.min", type="coefficients"))
as.numeric(predict(cv.lasso2, s = "lambda.min", type="coefficients"))
row.names(predict(cv.ridge, s = "lambda.min", type="coefficients"))
table2 = matrix(c(3061.1085  , 917.6644 ,  469.1203 , 1493.7920  ,-596.8185 ,  101.1146  ,-233.9494 ,-1941.7010,5818.3184,406.0477 ,-2610.9173 ,-3280.3364,30.611085,   9.176644 ,  4.691203,  14.937920 , -5.968185,1.011146,  -2.339494 ,-19.417010 , 58.183184,4.060477 ,-26.109173 ,-32.803364 ,2944.1794 ,  840.7957 ,439.4329 , 1383.8011 , -570.4896 ,  103.0900 , -247.3344, -1902.5122 , 3504.1926,2917.0274, -2466.6025,-3087.8805, 29.468977 ,  8.452095 ,  4.426609 , 13.896301 , -5.714296  , 1.030874  ,-2.465057, -19.041578 , 35.311387,28.895797 ,-24.755658, -30.930439, 3050.10123 ,  899.99200 ,  448.73319 , 1479.01403 , -590.47346  ,  98.38661 , -231.47830, -1932.26671,5744.73842  , 517.48958 ,-2602.77634, -3262.74584,28.5482360,   5.1809573  , 0.0000000 , 11.4923601 , -4.3710025 ,  0.2990248 , -1.7623000 ,-17.2240229,58.2129942,  10.3914228, -23.6076260, -29.0829508),ncol = 6)
rownames(table2) = c("(Intercept)","season2"   ,  "season3"  ,   "season4"   ,  "holiday1"   , "workingday1", "weathersit2","weathersit3" ,"temp"    ,    "atemp"    ,   "hum"      ,   "windspeed")
colnames(table2) = c("Linear", "Linear rescaled", "ridge", "ridge rescaled", "LASSO", "LASSO rescaled")
table2 %>% 
  knitr::kable() 
```

