---
title: "p8106_mtp"
author: "Abhishek Ajay"
date: "April 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ISLR)
library(glmnet)
library(caret)
library(corrplot)
library(plotmo)
library(splines)
library(lasso2)
library(ggplot2)
library(mgcv)
library(readr)
library(pdp)
library(earth)
```

##Introduction

Describe your data set. Provide proper motivation for your work.

What questions are you trying to answer?
How did you prepare and clean the data?

The following dataset has been obtained from the UCI Machine Learning Rpository. We are using the dataset that contains daily counts of bike rentals by both registered and casual members from the Capital Bikeshare System during the years 2011 and 2012. This dataset also contains weather data for each that's been extracted from the www.freemeteo.com while it uses the dc.gov page to extract information about the binary holiday column. 

Capital Bikeshare has over 350 stations in Washington, D.C., Alington, VA and MD, etc. 

We are trying to use this dataset to see if temperature, weather and type of day (working day, weekend or a holiday) have any effect on the amount of bike rentals on any given day. Simply put, we're trying to forecast bike rentals on any day. This can have a ton of implications in traffic management. Forecasting the amount of bikes on the road can give a better sense of travel time and allowing taxi services to provide more accurate "estimated time of arrival" estimates. 


```{r}
day = read.csv("./data/day.csv")
day_1 = read_csv("./data/bike_2015_q1.csv")
#head(day)

x = day %>% 
  select(-c(dteday, instant)) 
x = model.matrix(casual+registered+ cnt~., x)
corrplot(cor(x))
```


According to the above correlation plot, we decided to drop-- instant and intercept.

```{r}
day = day %>% 
  janitor::clean_names() %>% 
  mutate(season = as.factor(season), 
         yr = as.factor(yr), 
         mnth = as.factor(mnth), 
         holiday = as.factor(holiday), 
         weekday = as.factor(weekday), 
         workingday = as.factor(workingday),
         dteday = as.vector(dteday)
         ) %>% 
  select(-c(instant,dteday))
```

```{r}
x_1 = 
  day %>% 
  select(-c( casual, registered, cnt))

x_1 = model.matrix(~., x_1)
y = day$cnt
ridge.mod <-glmnet(x_1, y, alpha=0, lambda =exp(seq(-1, 10, length=100)))
ridge.mod$lambda

dim(coef(ridge.mod))

```

we dropped the dteday because it does is a factor variable with too many levels and it basically is the same as the index which does not tells us anything in this test. 

x_1: all independent variables
y: cnt---the response

```{r}
set.seed(2)
cv.ridge <-cv.glmnet(x_1, y,alpha = 0,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")
plot(cv.ridge)

plot_glmnet(ridge.mod, xvar = "rlambda")

best.lambda = cv.ridge$lambda.min
best.lambda
predict(ridge.mod, s = best.lambda, type="coefficients")
```


```{r}

cv.lasso <-cv.glmnet(x_1, y,alpha = 1,lambda =exp(seq(-1, 10, length=100)),type.measure = "mse")
cv.lasso$lambda.min
plot(cv.lasso)

plot_glmnet(cv.lasso$glmnet.fit)

predict(cv.lasso, s="lambda.min", type="coefficients")
```

? why is that lasso best lambda is 1 ?


```{r}
theme1 <-trellis.par.get()
theme1$plot.symbol$col <-rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <-rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <-rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x_1, y, plot = "scatter", labels =c("","Y"),type =c("p"), layout =c(4, 2))
```

actually this is pretty enough and straight forward to see some trends of these variables toward the bike usage. 





smooth splines for temp

```{r}
fit.ss <-smooth.spline(day$temp, day$cnt)
fit.ss$df

templims <-range(day$temp)
temp.grid <-seq(from = templims[1],to = templims[2], by = 0.0001)
length(temp.grid)

pred.ss <-predict(fit.ss,
                  x = temp.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        temp = temp.grid)

p <-ggplot(data = day,aes(x = temp, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

p+geom_line(aes(x = temp, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()



```

humidity

```{r}
fit.ss <-smooth.spline(day$hum, day$cnt)
fit.ss$df

humlims <-range(day$hum)
hum.grid <-seq(from = humlims[1],to = humlims[2], by = 0.001)
length(hum.grid)

pred.ss <-predict(fit.ss,
                  x = hum.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        hum = hum.grid)

p <-ggplot(data = day,aes(x = hum, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

p+geom_line(aes(x = hum, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()
```

pretty persuasive, since 0.5-0.75 is the best comfortable humidity



windspeed

```{r}
fit.ss <-smooth.spline(day$windspeed, day$cnt)
fit.ss$df

windspeedlims <-range(day$windspeed)
windspeed.grid <-seq(from = windspeedlims[1],to = windspeedlims[2], by = 0.001)
length(windspeed.grid)

pred.ss <-predict(fit.ss,
                  x = windspeed.grid)


pred.ss.df <-data.frame(pred = pred.ss$y,
                        windspeed = windspeed.grid)

p <-ggplot(data = day,aes(x = windspeed, y = cnt))+
  geom_point(color =rgb(.2, .4, .2, .5))

p+geom_line(aes(x = windspeed, y = pred), data = pred.ss.df,color =rgb(.8, .1, .1, 1))+ theme_bw()
```

windspeed ^  --  cnt V


GAM
```{r}
gam.m1 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed, data = day)


gam.m2 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + s(temp) + atemp + hum + windspeed, data = day) # s:temp
gam.m3 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + s(atemp) + hum + windspeed, data = day) # s:atemp
gam.m4 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + s(hum) + windspeed, data = day) # s:hum
gam.m5 = gam(cnt~ season + yr + mnth + holiday + weekday + workingday + weathersit + temp + atemp + hum + s(windspeed), data = day) # s:windspeed

anova(gam.m1, gam.m2, test = "F")
anova(gam.m1, gam.m3, test = "F")

par(mfrow=c(2,2))
plot(gam.m2)
plot(gam.m3)
plot(gam.m4)
plot(gam.m5)

#deleting the yr and mnth variables

gam.m1 = gam(cnt~ season +  holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed, data = day)


gam.m2 = gam(cnt~ season +  holiday + weekday + workingday + weathersit + s(temp) + atemp + hum + windspeed, data = day) # s:temp
gam.m3 = gam(cnt~ season +  holiday + weekday + workingday + weathersit + temp + s(atemp) + hum + windspeed, data = day) # s:atemp
gam.m4 = gam(cnt~ season +  holiday + weekday + workingday + weathersit + temp + atemp + s(hum) + windspeed, data = day) # s:hum
gam.m5 = gam(cnt~ season +  holiday + weekday + workingday + weathersit + temp + atemp + hum + s(windspeed), data = day) # s:windspeed

anova(gam.m1, gam.m2, test = "F")
anova(gam.m1, gam.m3, test = "F")

par(mfrow=c(2,2))
plot(gam.m2)
plot(gam.m3)
plot(gam.m4)
plot(gam.m5)

```

```{r}
#GCV to select df

ctrl1 <-trainControl(method = "cv", number = 10)
set.seed(2)
gam.fit <-train(x_1, y,method = "gam",tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),trControl = ctrl1)

gam.fit$bestTune
gam.fit$finalModel

x_2 = 
  day %>% 
  select(-c( casual, registered, cnt, yr, mnth))

x_2 = model.matrix(~., x_2)



set.seed(3)
gam.fit2 <-train(x_2, y,method = "gam",tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),trControl = ctrl1)

gam.fit2$bestTune
gam.fit2$finalModel
```

?The deviance is pretty huge of those ANOVA?

GCV score is too large. 

```{r}
#poisson regression for categorical terms, exclude yr and mnth


day.glm1 <- glm(cnt~season + holiday + weekday + workingday + weathersit, family=poisson, data=day)
summary(day.glm1)




```


```{r}
#mars
mars_grid <-expand.grid(degree = 1:2,nprune = 2:10)
set.seed(2)
mars.fit <-train(x_1, y,method = "earth",tuneGrid = mars_grid,trControl = ctrl1)
ggplot(mars.fit)
mars.fit$bestTune


coef(mars.fit$finalModel)
bwplot(resamples(list(mars = mars.fit,gam = gam.fit)), metric = "RMSE")
```

